---
title: "Hw1"
author: "Rachel Kaufman"
date: "2023-01-12"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(dplyr)
twitch_data <- read_csv("twitchdata-update.csv")
```

```{r}
library(janitor)

twitch_data <- clean_names(twitch_data)
colnames(twitch_data)

```

### **Question 1**
```{r}
View(twitch_data)
#looking at the data 
twitch_data %>% 
  select(average_viewers, followers)
#summarizing the data
twitch_data %>% 
  select(average_viewers, followers) %>% 
  summary()
```
In terms of what I notice, the range for minimum and maximum followers is incredibly wide distribution, and the mean is a lot higher than the median. As for the average viewers, it is noticeably less than the # of followers. The mean is also higher than the median, with a large distribution in range. Comparing their summary statistics does not tell me much about predictive power, but I can look at a scatter plot to help me out to look at their relationship :)


```{r}
## install.packages("hrbrthemes")
library(hrbrthemes)
ggplot(twitch_data, aes(average_viewers, followers)) +
  geom_point(size = 1, color = "blue") +
  labs(title = "Number of followers and Avg. number of views") 
```
There seems to be some sort of correlation, with a few wonky outliers. A lot of the data points are all in the bottom left, meaning they are stacked on each other and difficult to interpret, so log transform will give us a better understanding of the distribution. 

```{r}
ggplot(twitch_data, aes(average_viewers, followers)) +
  geom_point(size = 1, color = "blue") +
  labs(title = "Number of followers and Avg. number of views",
       x = "logged avg. viewers",
       y = "logged number of followers") +
  scale_x_log10() +
  scale_y_log10() 
```
Phew! Looks way better. There definitely seems to be a correlation. 

Adding the transformations as their own new columns to our data set:
```{r}
twitch_data <- twitch_data %>% 
  mutate(log_viewers = log10(average_viewers), 
         log_followers = log10(followers))
View(twitch_data)
```

### **Question 2**
```{r}
fit1 <- lm(formula = log_viewers~log_followers,
   data = twitch_data)
library(broom)
tidy(fit1)
##interpret your coefficients
1.1^{0.588} #1.1 -> is the 10%, this is just transforming the log into % for interpreting

```
This tells us with that when we see a 10% increase in number of followers, we see a 5.8% increase in average viewers. 

### **Question 3**
```{r}

##fitted values
pred_data <- augment(fit1)
glimpse(pred_data)

pred_data %>% 
  ggplot(aes(log_followers,
             log_viewers)) +
  geom_jitter(alpha = 0.4) +
  geom_line(aes(x = log_followers,
                y = .fitted),
            col = "orange") +
  theme_minimal() + 
  labs(title = "Followers and Average Viewership",
       subtitle = "Fitted model and Raw Data",
       x = "logged followers",
       y = "logged avg. viewers")
           
```
This shows us our predicted values given the line of best fit. 

```{r}
pred_data %>% 
  ggplot(aes(x = log_followers,
             y = .resid)) +
  geom_jitter(alpha = 0.4) +
  theme_minimal() + 
  labs(title = "Followers and Average Viewership",
       subtitle = "Residual model and Raw Data",
       x = "logged followers",
       y = "residuals")
```
So our residuals are pretty clustered [between -0.5 to 0.5] for their range. BUT! The range for our x-axis of logged followers is definitely skewed to the right. So, when logged followers is <5, we are in a sticky situation and predictions will likely suck. 

 ### **Question 4**
looking at language and avg. viewers. The goal of this is to work with a categorical variable to find predictive values of a continuous variable. Let's look at the raw data...
```{r}

#lets take a look at the variables we are interested in. 
twitch_data %>% 
  select(language, average_viewers)

#i want to be able to look at the # of different languages...
twitch_data %>% 
 count(language) %>% 
  arrange(desc(n))
```
So I have got 21 languages, with English having the highest count of 485. 

Lets use `summary()` to look at some descriptive statistics:
```{r}
twitch_data %>% 
  select(language, average_viewers) %>% 
  summary()

##visualizing language distribution
ggplot(twitch_data, aes(x = language)) +
  geom_bar() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

Now lets do a cute lil scatter plot which is really going to look like a bar chart bc language is a categorical variable. 
```{r}
ggplot(twitch_data, aes(language, average_viewers)) +
  geom_point() +
  labs(title = "Language and Average Viewers",
       y = "average number of viewers") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```
SHOCKER! They are all clustered around 0. I need to use the log transformed variable I set up in one of the earlier questions.  

```{r}
ggplot(twitch_data, aes(language, log_viewers)) +
  geom_point() +
  labs(title = "Language and Average Viewers",
       y = "logged average number of viewers") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```
I don't really see much, but I am not sure what would be helpful in this instance. I think any hunch that I would have is basically that Russian and English seem to have some upward bound-esk residuals that I think will definitely pull the model in in terms of these two categories having a larger relative magnitude in comparison to other languages like Thai and Czech, where their spread is smaller in range and both stay below 4 for the logged average number of viewers.



### **Question 5**
```{r}
#setting English as reference category
twitch_data <- twitch_data %>% 
  mutate(language = as.factor(language), 
         language = relevel(language, ref = "English"))
 
fit2 <-  
  lm(data = twitch_data,
     formula = log_viewers ~ language)


pred_data2 <- augment(fit2)
pred_data2

tidy(fit2)
View(pred_data2)
```
In terms of what the coefficients tell us, yes Nico your hunch was correct that English has quite the magnitude in comparison to the other languages. For example, when we look at Finnish as a category we can see the decrease by 0.453 log units of viewers in comparison to English viewership. In terms of positive coefficients, we have both Russian, Japanese and Spanish with positive coefficients. This means that in comparison to English they actually saw an increase in logged viewership. So, this kinda says to Nico's hunch wellllllll you were close but not that close because if these are in reference to english that means that they predict more viewership.

### **Question 6**
```{r}
pred_data2 %>% 
  ggplot(aes(x = language,
             y = .resid)) +
  geom_jitter(alpha = 0.4) +
  theme_minimal() + 
  labs(title = "Languages and Average Viewership",
       subtitle = "Residuals from model",
       x = "Languages",
       y = "Residuals") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) 


## looking at my little residuals
pred_data2 %>% filter(.resid > 1.5)

```
So, yes there were quite a few our model missed!BUT the two largest are for English and Russian with logged viewers of 5.17 and 5.10




