---
title: "Hw1"
author: "Rachel Kaufman"
date: "2023-01-12"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(dplyr)
twitch_data <- read_csv("twitchdata-update.csv")
```

```{r}
library(janitor)

twitch_data <- clean_names(twitch_data)
colnames(twitch_data)

```

Q1
```{r}
View(twitch_data)
#looking at the data 
twitch_data %>% 
  select(average_viewers, followers)
#summarizing the data
twitch_data %>% 
  select(average_viewers, followers) %>% 
  summary()
```
In terms of what I notice, the range for minimum and maximum followers is incredibly wide distribution, and the mean is a lot higher than the median. As for the average viewers, it is noticeably less than the # of followers. The mean is also higher than the median, with a large distribution in range. Comparing their summary statistics does not tell me much about predictive power, but I can look at a scatter plot to help me out to look at their relationship :)


```{r}
install.packages("hrbrthemes")
library(hrbrthemes)
ggplot(twitch_data, aes(average_viewers, followers)) +
  geom_point(size = 1, color = "blue") +
  labs(title = "Number of followers and Avg. number of views") 
```
There seems to be some sort of correlation, with a few wonky outliers. A lot of the data points are all in the bottom left, meaning they are stacked on each other and difficult to interpret, so log transform will give us a better understanding of the distribution. 

```{r}
ggplot(twitch_data, aes(average_viewers, followers)) +
  geom_point(size = 1, color = "blue") +
  labs(title = "Number of followers and Avg. number of views",
       x = "logged avg. viewers",
       y = "logged number of followers") +
  scale_x_log10() +
  scale_y_log10() 
```
Phew! Looks way better. There definitely seems to be a correlation. 

Adding the transformations as their own new columns to our data set:
```{r}
twitch_data <- twitch_data %>% 
  mutate(log_viewers = log10(average_viewers), 
         log_followers = log10(followers))
View(twitch_data)
```

### **Q2**
```{r}
fit1 <- lm(formula = log_viewers~log_followers,
   data = twitch_data)
library(broom)
tidy(fit1)
##interpret your coefficients
1.1^{0.588} #1.1 -> is the 10%, this is just transforming the log into % for interpreting

```
This tells us with that when we see a 10% increase in number of followers, we see a 5.8% increase in average viewers

### **Q3**
```{r}

##fitted values
pred_data <- augment(fit1)
glimpse(pred_data)

pred_data %>% 
  ggplot(aes(log_followers,
             log_viewers)) +
  geom_jitter(alpha = 0.4) +
  geom_line(aes(x = log_followers,
                y = .fitted),
            col = "orange") +
  theme_minimal() + 
  labs(title = "Followers and Average Viewership",
       subtitle = "Fitted model and Raw Data",
       x = "logged followers",
       y = "logged avg. viewers")
           
```
This shows us our predicted values given the line of bestfit. 

```{r}

pred_data %>% 
  ggplot(aes(log_followers,
             .resid)) +
  geom_jitter(alpha = 0.4) +
  theme_minimal() + 
  labs(title = "Followers and Average Viewership",
       subtitle = "Residual model and Raw Data",
       x = "logged followers",
       y = "residuals")
```
So our residuals are pretty clustered [between -0.5 to 0.5] for their range. BUT! The range for our x-axis of logged followers is definitely skewed to the right. So, when logged followers is <5, we are in a sticky situation and predictions will likely suck. 

Q4
looking at language and avg. viewers. The goal of this is to work with a categorical variable to find predictive values of a continuous variable. Let's look at the raw data...
```{r}

#lets take a look at the variables we are interested in. 
twitch_data %>% 
  select(language, average_viewers)

#i want to be able to look at the # of different languages...
twitch_data %>% 
 count(language) %>% 
  arrange(desc(n))
```
So I have got 21 languages, with English having the highest count of 485. 

Lets use `summary()` to look at some descriptive statistics:
```{r}
twitch_data %>% 
  select(language, average_viewers) %>% 
  summary()

##visualizing language distribution
ggplot(twitch_data, aes(x = language)) +
  geom_bar() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

Now lets do a cute lil scatter plot which is really going to look like a bar chart bc language is a categorical variable. 
```{r}
ggplot(twitch_data, aes(language, average_viewers)) +
  geom_point() +
  labs(title = "Language and Average Viewers",
       y = "average number of viewers") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

SHOCKER! They are all clustered around 0. I need to use the log transformeed variable I set up in 

```{r}
ggplot(twitch_data, aes(language, log_viewers)) +
  geom_point() +
  labs(title = "Language and Average Viewers",
       y = "logged average number of viewers") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```
I don't really see much, but I am not sure what would be helpful in this instance. 

### **Q4**
```{r}
#setting English as reference category
twitch_data <- twitch_data %>% 
  mutate(language = as.factor(language), 
         language = relevel(language, ref = "English"))

##changing the values of the languages to #'s
twitch_data <- twitch_data %>% 
  mutate(language_factor = case_when(
         language == "English" ~ 1,
         language == "Arabic" ~ 2,
         language == "Chinese" ~ 3,
         language == "Czech" ~ 4,
         language == "Finnish" ~ 5,
         language == "French" ~ 6,
         language == "German" ~ 7,
         language == "Greek" ~ 8,
         language == "Hungarian" ~ 9,
         language == "Italian" ~ 10,
         language == "Japanese" ~ 11,
         language == "Korean" ~ 12,
         language == "Polish" ~ 13,
         language == "Portuguese" ~ 14,
         language == "Russian" ~ 15,
         language == "Slovak" ~ 16,
         language == "Spanish" ~ 17,
         language == "Swedish" ~ 18,
         language == "Thai" ~ 19,
         language == "Turkish" ~ 20,
         language == "Other" ~ 21)
         )
 
fit2 <-  
  lm(data = twitch_data,
     formula = log_viewers ~ language_factor)


pred_data2 <- augment(fit2)
fit2
View(pred_data2)
```

```{r}
pred_data2 %>% 
  ggplot(aes(language_factor,
             log_viewers)) +
  geom_jitter(alpha = 0.4) +
  geom_line(aes(x = language_factor,
                y = .fitted),
            col = "orange") +
  theme_minimal() + 
  labs(title = "Languages and Average Viewership",
       subtitle = "Fitted model and Raw Data",
       x = "Languages",
       y = "logged avg. viewers")
           
```

### **Question 5**
```{r}

```

### **Question 6**
```{r}

```


