---
title: "hw3_review"
author: "Rachel Kaufman"
date: "2023-01-18"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
##install.packages("moderndive")
library(moderndive)
##install.packages("bayesrules")
theme_set(theme_minimal())

data(bikes, package = "bayesrules")
glimpse(bikes)
```
Well, the way it is presented via glimpse it tells us there are 500 rows and 12 columns. 500 rows means no. of observations. The dates covered from the dataset are
```{r}
head(bikes$date)
tail(bikes$date)
```
Dates represent Jan 1 2011 to December 31st 2012. 
what does each row represent?
what dates does the dataset cover?
what is the highest observed ridership in the dataset?
```{r}
max(bikes$rides)
max(bikes$windspeed)
```
So, the highest observed ridership is 6946 rides and the highest windspeed is 34 mph


### **Question 1**
```{r}
bikes %>% 
  select(rides, temp_feel, windspeed) %>% 
  cor(use = "complete.obs")
```
The correlation between what the temperature feels like and number of rides is 0.582 and then for windspeed and rides the correlation is -0.19. So, temp feel has a higher correlation. 

### **Question 2**
Using the approximation that a mile is equal to 1.61 kilometers, convert windspeed to kilometers per hour. Call the new variable wind_kph and add it to the bikes data frame. What is the correlation between wind speed in MPH and wind speed in KPH? Explain why in enough detail that I know you understand.
```{r}
bikes <- bikes %>%
  mutate(wind_KPH = windspeed*1/1.61)
View(bikes)

bikes %>% 
  select(wind_KPH, windspeed) %>% 
  cor(use = "complete.obs")
```
So! It should be 1... ya know because this is a conversion... it should not change anything. By using a conversion (by using proportions) they have to be equal to it... literal math laws

### **Question 3**
```{r}
kph_model <- lm(data = bikes,
                formula = rides~ wind_KPH)
broom :: tidy(kph_model)

mph_model  <- lm(data = bikes,
                formula = rides~ windspeed)
broom :: tidy(mph_model)

```
Okay, so basically if looking at mph its that when we see a unit change in windspeed we see a -55.5 unit change for # of rides. And similarly, wind speed in kph, is when we see a unit increase in KM per hour there is a -89.4 unit change for # of rides... again, this is a conversion so uh this sort of makes sense. I think we see a larger slope for the kph model because well if it is 1.61 miles to 1 km, then the distribution these points is going to be shorter in range, because this measure is going to quite literally cover a spread of less numbers in comparison to mph model as a counterpart. the std error being different also supports this conclusion. these are measures of variance babeyyyy


### **Question 4**
Using the models from above, tell me what the predicted ridership would be if the wind is blowing at 20 KPH. What would the predicted ridership be if the wind is blowing at 20 MPH?
```{r}
##kph
4205 + -89.4*20
#mph
4205 + -55.5*20
```
The predicted ridership for kph is 2417; for mph it is 3095 rides. 

### **Question 5**
Creating a new variable called temp_c that is a conversion of temp_feel to Celsius and add it to the bikes dataframe. 

With this new variable, estimate a multiple regression that predicts rides as an additive function of wind_kph and temp_c (that is, no interaction). Interpret both of the slope coefficients using the following language:
```{r}
bikes <- bikes %>% 
  mutate(temp_c = 5/9*(temp_feel - 32))

 multipleRegression_model <- lm(data = bikes,
                              formula = rides ~ wind_KPH + temp_c)
 
broom :: tidy(multipleRegression_model)
```
For a one unit increase in temp_c, the model expects a 142 increase in rides.
For a one unit increase in wind_KPH, the model expects a 51 decrease in rides.
 
### **Question 6**
```{r}
pred_df <- tibble(
  situation = 1:3,
  temp_c = c(25, 15, 10),
  wind_kph = c(15, 5, 40)
)
pred_df

942 + -51.4*15 + 25*143
942 + -51.4*5 + 15*15
942 + -51.4*40 + 10*143
```





















